---
title: "Alkylbenzene PCA"
author: "BH"
date: "2023-01-03"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code written and executed by **Blake Hunnie** (hunnieb@myumanitoba.ca)
This R-markdown file is intended for the reproducibility of manipulating and presenting data from the BIOS site.
Most of the code included below comes from the **tidyverse**, rather than solely **Base-R**.

# All necessary data & information must be loaded into R:
## First, the packages included in my typical workflows are loaded:

```{r loading packages}

library(tidyverse)
library(janitor)
library(here)
library(gt)
library(broom) 
library(fs)
library(stats)
library(ggrepel)
library(patchwork)
library(ggfortify)
library(knitr)

```


```{r loading dataset}

alkylbenzene_dataset <- read_csv(here("data", "tidy_individual_AB_pca.csv")) #loading in the csv.file dataset

```


```{r removing categorical variables}

ab_dataset_continuous <- alkylbenzene_dataset %>% #assigning the dataset to an object
  select(-site, -sample_id, -sample_depth, -sample_type) #removing any non-continuous data columns
ab_dataset_continuous #calling the object

```


```{r running the PCA}

ab_pca <- ab_dataset_continuous %>% #assigning the continuous dataset to an object
  prcomp(center = TRUE, scale = TRUE) #performing a PCA
summary(ab_pca) #summarizing the results from the object

```


```{r binding PC1 and PC2 to the original dataset}

combined_data <- cbind(alkylbenzene_dataset, ab_pca$x[,1:3]) #combining the first 3 columns of the PCA results to the initial table
combined_data #calling the combined dataset

```

##The next three chunks of code are nearly identical; They're meant to produce figures representing the results of PC1xPC2, PC1xPC3, and PC2xPC3 (only the first chunk will have comments)

```{r PC1xPC2}

PC1_PC2_plot <- combined_data %>% #assigning the combined dataset to an object
  ggplot() + #creating a ggplot item
  stat_ellipse(aes(x = PC1, y = PC2, colour = site, fill = site), geom = "polygon", alpha = 0.5) + #creating a 95% confidence interval ellipse
  geom_point(aes(x = PC1, y = PC2, colour = site, fill = site), shape = 21, colour = "black") + #plotting the PC1xPC2 results
  geom_text_repel(aes(x = PC1, y = PC2, label = sample_id), size = 3) + #preventing text from overlapping on the plot
  geom_hline(aes(yintercept = 0), linetype = "dashed", alpha = 0.5) + #creating a dashed horizontal line on y=0
  geom_vline(aes(xintercept = 0), linetype = "dashed", alpha = 0.5) + #creating a dashed vertical line on x=0
  labs(subtitle = "A") + #adding a subtitle
  theme_classic() #assigning the classic tidyverse theme
PC1_PC2_plot #calling the plot
  
ggsave(here("figures", "pc1_pc2.pdf"), PC1_PC2_plot, #saving the plot as a pdf. file
       width = 190, height = 160, units = "mm") #setting the size parameters of the plot

```


```{r PC1xPC3}

PC1_PC3_plot <- combined_data %>%
  ggplot() +
  stat_ellipse(aes(x = PC1, y = PC3, colour = site, fill = site), geom = "polygon", alpha = 0.5) +
  geom_point(aes(x = PC1, y = PC3, colour = site, fill = site), shape = 21, colour = "black") +
  geom_text_repel(aes(x = PC1, y = PC3, label = sample_id), size = 3) +
  geom_hline(aes(yintercept = 0), linetype = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = 0), linetype = "dashed", alpha = 0.5) +
  labs(subtitle = "B") +
  theme_classic() 
PC1_PC3_plot
  
ggsave(here("figures", "pc1_pc3.pdf"), PC1_PC3_plot,
       width = 190, height = 120, units = "mm") 

```


```{r PC2xPC3}

PC2_PC3_plot <- combined_data %>%
  ggplot() +
  stat_ellipse(aes(x = PC2, y = PC3, colour = site, fill = site), geom = "polygon", alpha = 0.5) +
  geom_point(aes(x = PC2, y = PC3, colour = site, fill = site), shape = 21, colour = "black") +
  geom_text_repel(aes(x = PC2, y = PC3, label = sample_id), size = 3) +
  geom_hline(aes(yintercept = 0), linetype = "dashed", alpha = 0.5) +
  geom_vline(aes(xintercept = 0), linetype = "dashed", alpha = 0.5) +
  labs(subtitle = "C") +
  theme_classic() 
PC2_PC3_plot
  
ggsave(here("figures", "pc2_pc3.pdf"), PC2_PC3_plot,
       width = 190, height = 120, units = "mm") 

```

##This next section serves to combine the three above plots into a single file

```{r patching the plots together}

full_plot <- #creating an object
  PC1_PC2_plot + #adding the first plot to the object
  PC1_PC3_plot + #adding the second plot to the object
  PC2_PC3_plot + #adding the third plot to the object
  plot_layout(ncol = 1, guides = 'collect') & theme(legend.position = 'bottom') #assigning the layout of the three plots
full_plot #calling the plot

ggsave(here("figures", "full_plot.pdf"), full_plot, #saving the plot as a pdf. file
       width = 190, height = 290, units = "mm") #assigning the size parameters to the plot


```



