---
title: "Percent Residual Comparison"
author: "BH"
date: "2023-01-06"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code written and executed by **Blake Hunnie** (hunnieb@myumanitoba.ca)
This R-markdown file is intended for the reproducibility of manipulating and presenting data from the BIOS site.
Most of the code included below comes from the **tidyverse**, rather than solely **Base-R**.

# All necessary data & information must be loaded into R:
## First, the packages included in my typical workflows are loaded:

```{r loading packages}

library(tidyverse)
library(janitor)
library(here)
library(gt)
library(broom) 
library(fs)
library(viridis)

```


```{r loading percent residual data}

percent_residual_data <- read_csv(here("data", "tidy_percent_residual_comparison.csv")) %>% #loading in the csv.file dataset
  clean_names() %>% #altering the column names in accordance to the tidyverse
  rename(nC18 = n_c18, nC30 = n_c30) %>% #renaming columns 
  pivot_longer(cols = c(nC18, phytane, nC30), names_to = "parameter", values_to = "measure") #converting numerous columns into fewer
percent_residual_data #calling the dataset

```


```{r plotting the data}

percent_residual_plot <- percent_residual_data %>% #assigning the dataset to an object
  mutate(parameter = factor(parameter, levels = c("nC18", "phytane", "nC30"))) %>% #assigning the order of the variables
  ggplot() + #creating the ggplot item
  facet_wrap(~sample_year, ncol = 2) + #creating a two-windowed plot
  geom_col(aes(x = sample_id, y = measure, fill = parameter), colour = "black", position = "dodge") + #creating bar plots
  labs(y = "Percent Residual (%)") + #labelling the y-axis
  scale_x_discrete(limits = c("B11-I 0-2", "B11-I 5-10", "T1 0-2", "T2 0-2", "IMC-c 0-2", "IMC-c 5-10")) + #assigning the order of the samples
  scale_y_continuous(breaks = seq(0, 100, by = 20)) + #setting the y-axis tick breaks
  theme_minimal() + #applying the tidyverse minimal theme to the plot
  scale_fill_viridis(discrete = TRUE) + #changing the colour palette
  theme(axis.title.x = element_blank(), #removing the x-axis title
        axis.text.x = element_text(size = 7), #changing the size of the x-axis text
        legend.title = element_blank(), #removing the title from the legend
        axis.ticks.x = element_line(size = 0.25), #changing the size of the x-axis tick marks
        panel.grid.major.x = element_blank(), #removing the panelling
        legend.position = "bottom") #altering the position of the legend
percent_residual_plot #calling the plot

ggsave(here("figures", "percent_residual_comparison.pdf"), percent_residual_plot, #saving the plot as a pdf. file
       width = 190, height = 120, units = "mm") #setting the size parameters of the save file

```


