---
title: "Alkylcycloalkane Data"
author: "BH"
date: "2023-01-02"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code written and executed by **Blake Hunnie** (hunnieb@myumanitoba.ca)
This R-markdown file is intended for the reproducibility of manipulating and presenting data from the BIOS site.
Most of the code included below comes from the **tidyverse**, rather than solely **Base-R**.

# All necessary data & information must be loaded into R:
## First, the packages included in my typical workflows are loaded:

```{r loading packages}

library(tidyverse)
library(janitor)
library(here)
library(gt)
library(broom) 
library(fs)
library(viridis)

```



```{r loading dataset}

alkylcycloalkane_data <- read_csv(here("data", "tidy_individual_aca_percentages.csv")) %>% #loading in the csv.file dataset
  clean_names() %>% #altering column names in accordance with the tidyverse
  pivot_longer(cols = c(n_hexylcyclohexane, n_heptylcyclohexane, n_octylcyclohexane, n_nonylcyclohexane, n_decylcyclohexane, n_undecylcyclohexane, n_dodecylcyclohexane, n_tridecylcyclohexane, n_tetradecylcyclohexane), names_to = "parameter", values_to = "measure") #taking many columns and condensing them into fewer, longer columns
alkylcycloalkane_data #calling the dataset

```



```{r nesting data}

nested_data <- alkylcycloalkane_data %>% #assigning the dataset to an object
  group_by(site, sample_depth, sample_type) %>% #grouping the data based on 3 variables: site, sample depth, and sample type
  nest() #nesting the data
nested_data #calling the nested data

```


```{r creating nested plots}

nested_plots <- nested_data %>% #assigning the nested data to an object
  mutate(plot = pmap(list(site, sample_type, sample_depth, data), #creating individual objects to automatically generate plots from
                     ~ ggplot() + #creating a ggplot item
                       geom_col(data = ..4, aes(x = parameter, y = measure, fill = sample_id), colour = "black", position = "dodge") + #sequentially creating bar plots from the nested data
                       labs(y = "Percent of total (%)") + #labelling the y-axis
                       scale_x_discrete(limits = c("n_hexylcyclohexane", "n_heptylcyclohexane", "n_octylcyclohexane", "n_nonylcyclohexane", "n_decylcyclohexane", "n_undecylcyclohexane", "n_dodecylcyclohexane", "n_tridecylcyclohexane", "n_tetradecylcyclohexane")) + #ordering the observations on the x-axis
                       theme_minimal() + #applying the tidyverse minimal theme to the plots
                       scale_fill_viridis(discrete = TRUE) + #changing the colour palette
                       theme(axis.title.x = element_blank(), #removing the title from the x-axis
                             axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.9), #changing the angle of the x-axis text
                             legend.title = element_blank(), #removing the title from the legend
                             axis.ticks.x = element_line(size = 0.25), #changing the size of the x-axis ticks
                             panel.grid.major.x = element_blank(), #removing the panelling
                             legend.position = "bottom"), #moving the position of the legend
                     labs(title = str_c(..1, ..2, ..3, "sediments", sep = ", "))), #labelling the plots based on the sequential variables being observed
         filename = str_c(site, sample_type, sample_depth, "plot.pdf", sep = "_")) #saving the filenames based on these variables
nested_plots #calling the list of plots

```



```{r saving the individual plots}

nested_plots_save <- nested_plots %>% #assigning the list of plots to an object
  ungroup() %>%  #ungrouping the data to save individual files
  select(filename, plot) %>% #selecting the two parameters to save the files from
  pwalk(ggsave, path =  here("figures"), width = 190, height = 120, units = "mm") #sequentially saving each file

```

##This next section is dedicated to creating individual plots from the control sample datasets 

```{r loading in mean control dataset}

aca_control <- read_csv(here("data", "controls_alkylcycloalkanes.csv")) %>% #importing the csv.file dataset
  clean_names() %>% #renaing the columns in accordance to the tidycerse
  pivot_longer(cols = c(n_hexylcyclohexane, n_heptylcyclohexane, n_octylcyclohexane, n_nonylcyclohexane, n_decylcyclohexane, n_undecylcyclohexane, n_dodecylcyclohexane, n_tridecylcyclohexane, n_tetradecylcyclohexane), names_to = "parameter", values_to = "measure") #onsolidating many columns into fewer columns
aca_control #calling the data

```


```{r creating control plot}

control_plot <- aca_control %>%  #assigning the control dataset to an object
  mutate(across(sample_id, factor, levels=c("surface control","subsurface control"))) %>% #separating the results based on sample depth
  ggplot() + #creating a ggplot item
  geom_col(aes(x = parameter, y = measure, fill = sample_id), colour = "black", position = "dodge") + #creating bar plots
  labs(y = "Percent of total (%)") + #labelling the y-axis
  scale_x_discrete(limits = c("n_hexylcyclohexane", "n_heptylcyclohexane", "n_octylcyclohexane", "n_nonylcyclohexane", "n_decylcyclohexane", "n_undecylcyclohexane", "n_dodecylcyclohexane", "n_tridecylcyclohexane", "n_tetradecylcyclohexane")) + #assigning the order of the observations on the x-axis
  theme_minimal() + #applying the minimal tidyverse theme to the plot
  scale_fill_viridis(discrete = TRUE) + #changing the colour palette 
  theme(axis.title.x = element_blank(), #removing the title from the x-axis
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.9), #changing the angle of the x-axis text
        legend.title = element_blank(), #removing the title from the legend
        axis.ticks.x = element_line(size = 0.25), #changing the size of the x-axis tick marks
        panel.grid.major.x = element_blank(), #removing the panelling
        legend.position = "bottom") #changing the position of the legend
control_plot #calling the plot

ggsave(here("figures", "control_plot.pdf"), control_plot, #saving the plot as a pdf file
       width = 190, height = 120, units = "mm") #csetting the size parameters of the plot

```

##Finally, this section does the same as the control samples, but for the technical mixture (refer to comments in the above section)

```{r loading in technical mixture dataset}

lagomedio_data <- read_csv(here("data", "tidy_lagomedio_aca_percentages.csv")) %>%
  clean_names() %>%
   pivot_longer(cols = c(n_hexylcyclohexane, n_heptylcyclohexane, n_octylcyclohexane, n_nonylcyclohexane, n_decylcyclohexane, n_undecylcyclohexane, n_dodecylcyclohexane, n_tridecylcyclohexane, n_tetradecylcyclohexane), names_to = "parameter", values_to = "measure")

```


```{r generating technical mixture plot}

lagomedio_plot <- lagomedio_data %>%
  ggplot() +
  geom_col(aes(x = parameter, y = measure, fill = sample_id), colour = "black", position = "dodge") +
  labs(y = "Percent of total (%)") +
  scale_x_discrete(limits = c("n_hexylcyclohexane", "n_heptylcyclohexane", "n_octylcyclohexane", "n_nonylcyclohexane", "n_decylcyclohexane", "n_undecylcyclohexane", "n_dodecylcyclohexane", "n_tridecylcyclohexane", "n_tetradecylcyclohexane")) +
  theme_minimal() +
  #scale_fill_viridis(discrete = TRUE) +
  theme(axis.title.x = element_blank(),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 0.9),
        legend.title = element_blank(),
        axis.ticks.x = element_line(size = 0.25),
        panel.grid.major.x = element_blank(),
        legend.position = "bottom")
lagomedio_plot

ggsave(here("figures", "lagomedio_plot.pdf"), lagomedio_plot,
       width = 190, height = 120, units = "mm") 

```




