---
title: "N-alkane PCA"
author: "BH"
date: "2023-01-10"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Code written and executed by **Blake Hunnie** (hunnieb@myumanitoba.ca)
This R-markdown file is intended for the reproducibility of manipulating and presenting data from the BIOS site.
Most of the code included below comes from the **tidyverse**, rather than solely **Base-R**.

# All necessary data & information must be loaded into R:
## First, the packages included in my typical workflows are loaded:

```{r loading packages}

library(tidyverse)
library(janitor)
library(here)
library(gt)
library(broom) 
library(fs)
library(stats)
library(ggrepel)
library(patchwork)
library(ggfortify)
library(knitr)
library(viridis)

```


```{r loading in dataset}

alkane_dataset <- read_csv(here("data", "tidy_individual_alkanes_pca.csv")) #loading in the csv.file dataset

```


```{r performing the PCA}

nalk_pca <- alkane_dataset %>% #assigning the loaded dataset to an object
  nest() %>% #nesting the data
  mutate(pca = map(data, ~prcomp(.x %>% select(-site, -sample_id), center = TRUE, scale = TRUE)), #performing a PCA while distinguishing results based on the sites and sample IDs
         pca_aug = map2(pca, data, ~augment(.x, data = .y))) #adding augmented data from the PCA to the dataset 
nalk_pca #calling the object

```


```{r plotting the PCA}

test_plot <- nalk_pca %>% #assigning the PCA results to an object
  mutate(pca_graph = map2(.x = pca, .y = data, 
                          ~autoplot(.x, loadings = TRUE, loadings.label = TRUE, loadings.label.repel = TRUE, loadings.colour = "black", loadings.label.colour = "black",
                                    data = .y, label = TRUE, label.label = "sample_id", label.repel = TRUE, colour = "site", face = "bold") + #plotting the results of the PCA
                            geom_hline(aes(yintercept = 0), linetype = "dashed", alpha = 0.5) + #adding a dashed horizontal line on y=0
                            geom_vline(aes(xintercept = 0), linetype = "dashed", alpha = 0.5) + #adding a dashed vertical line on x=0
                            theme_classic() + #applying the classic tidyverse theme to the plot
                            theme(legend.title = element_blank(), #removing the legend title
                                  legend.position = "bottom"))) #changing the position of the legend
test_plot$pca_graph #calling the PCA plot

#The plot can be saved by right-clicking the image and selecting "save image as."

```


